MCR_Algae_Data_Cleaning

### 2024-07-03 mcr_algal_data_cleaning_final.Rmd
adding in QA/QC step
Julianna reviewed cleaning code--looks great! :)

---
title: "MCR_Algae_Data_Cleaning"
author: "Noam Altman-Kurosaki and Lauren Enright"
date: "2024-05-03"
output: html_document
---

@@

rm(list=ls()) # clean up
library(knitr)
library(dplyr)
library(plyr)
library(car)
library(reshape)
library(reshape2)
library(tidyr)
library(effects)
library(devtools)
opts_chunk$set(comment="  ",
               collapse=TRUE, 
               echo=FALSE,
               #fig.asp=1/gr,
               fig.height=8,
               fig.width = 10,
               dev="png",
               warning=TRUE
               )
opts_knit$set(eval.after = "fig.cap")

@@

```{r data cleaning}
library(here)

data <- read.csv(here("data", "MCR_LTER_Annual_Survey_Benthic_Cover_20231211.csv"), stringsAsFactors = F) # replacements don't work when strings are factors
str(data)
levels(as.factor(data$Taxonomy_Substrate_Functional_Group)) # 86 levels
data$Site_habitat <- as.factor( paste(data$Site, data$Habitat))

@@

data %>% 
  filter(Taxonomy_Substrate_Functional_Group == "No data")

#99 observations that do not equal 100

@@

@@

@@

#how many are no data (-1)
comb_wide_quad_sum %>%
  dplyr::filter(row_sums == -1)

@@

mistakes_wide_quad_level <- comb_wide_quad_sum %>%
  dplyr::filter(row_sums != 100) %>%
  dplyr::filter(row_sums != -1)

comb_wide_quad_sum %>%
 dplyr::filter(row_sums != 100) %>% dplyr::filter(Year == 2020)

@@

#should be 19750 obs

@@

@@

>>>>

---
title: "MCR_Algae_Data_Cleaning"
author: "Noam Altman-Kurosaki and Lauren Enright"
date: "2024-05-03"
output: html_document
editor_options: 
  chunk_output_type: console
---

@@

rm(list=ls()) # clean up
library(tidyverse)
library(knitr)
library(dplyr)
library(plyr)
library(car)
library(reshape)
library(reshape2)
library(tidyr)
library(effects)
library(devtools)
opts_chunk$set(comment="  ",
               collapse=TRUE, 
               echo=FALSE,
               #fig.asp=1/gr,
               fig.height=8,
               fig.width = 10,
               dev="png",
               warning=TRUE
               )
opts_knit$set(eval.after = "fig.cap") 
library(here)

@@

```{r data cleaning}
data <- read.csv(here("data", "MCR_LTER_Annual_Survey_Benthic_Cover_20231211.csv"), stringsAsFactors = F) # replacements don't work when strings are factors

str(data)

levels(as.factor(data$Taxonomy_Substrate_Functional_Group)) # 86 levels
data$Site_habitat <- as.factor( paste(data$Site, data$Habitat))

@@

# can see that there are 99 observations that do not equal 100
dim(non100_quad_level)

# can look at which ones have no data 
data %>% 
  filter(Taxonomy_Substrate_Functional_Group == "No data")

@@

Just to check, Julianna is going to try this a different way (no mean function--that shouldn't be necessary, right?). Can delete this next chunk if you want:

```{r}
data %>% 
  # doing this Date removal because there is actually one that has multiple dates for the quadrat... must be a typo? If you don't filter date, get an extra row
  select(-Date) %>% 
  pivot_wider(names_from = Taxonomy_Substrate_Functional_Group, 
              values_from = Percent_Cover) %>% 
  replace(is.na(.), 0) -> testWide

# check
summary(testWide) 
summary(comb_wide) # eyeball perusal suggests they look identical
```

@@

comb_wide_quad_sum %>%
  dplyr::group_by(Year, Site, Habitat) %>%
  dplyr::summarise(count = n()) # 50 quadrats per habitat/year/site (this is correct)

@@

#how many are no data (-1)
comb_wide_quad_sum %>%
  dplyr::filter(row_sums == -1) %>% dim()

@@

# grab the mistake quadrats 
mistakes_wide_quad_level <- comb_wide_quad_sum %>%
  dplyr::filter(row_sums != 100) %>%
  dplyr::filter(row_sums != -1)

comb_wide_quad_sum %>%
  dplyr::filter(row_sums != 100) %>% 
  dplyr::filter(Year == 2020) %>% dim()

@@

#should be 19750 obs
dim(wide_alldata_filtered)

@@

# Julianna is adding this code here IF we want it:
wide_alldata_filtered_cleaned_column_names %>% 
  mutate(location = gsub(" ", "_", str_to_lower(location)),
         site = gsub(" ", "_", str_to_lower(site)),
         habitat = gsub(" ", "_", str_to_lower(habitat)),
         site_habitat = gsub(" ", "_", str_to_lower(site_habitat)),
         ) # would want to save this then do it for the algae dataset as well

@@

#should be 200 per year/site combo
